---
title: "ESM 244 - Lab 1"
author: "Kerry Nixon"
date: "1/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
```


### Read in SF trees data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Basic wrangling reminders

Refresh some skills for data wranginling and summary statistics using functions in the `dplyr` package.

find the top 5 highest observations of trees by legal_status, do some wrangling and make a graph.

count() can replace group_by summaraize. You can caount by multple variables seperated by commas. `count(variable_1, variable_2)`

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% # count by listed variable or variables
  drop_na(legal_status) %>% # remove rows with na in listed variable
  rename(tree_count = n) %>% # rename(new name = old name)
  relocate(tree_count) %>% # move tree_count to first column position
  slice_max(tree_count, n=5) # select the top 5 highest values
```

Make a graph of top 5 observations bylegal status
```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col() +
  labs(x = "Legal Status", y = "Tree Count") +
  coord_flip() +
  theme_minimal()
  
```

### A few more data wrangling refresher examples

Only want to keep observations (rows) for Blackwood Acacia Trees

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% #will look within a variable we specify adn see if a string is detected anywhere in that variable
  select(legal_status, date, latitude, longitude)

ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude)) +
  geom_point()
```

### tidyr::separate() and unite() functions

Useful for combining and separating columns

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```

Example of tidyr::unite()

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool_")
```


### Make some actual maps of blackwood acacia trees in SF

`st_as_sf()` to convret latitude and longitude to spatial coordinates. 

```{r}
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(longitude, latitude) %>% 
  st_as_sf(coords = c("longitude","latitude"))

st_crs(blackwood_acacia_sp) = 4326

ggplot(data = blackwood_acacia_sp) +
  geom_sf(color = "darkgreen")
```

Read in SF roads shapefile:

```{r}
sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

st_transform(sf_map, 4326)

ggplot(data = sf_map) +
  geom_sf()
```

COmbine blackwood acacia tree observcations & SF roads map:

```{r}
ggplot() +
  geom_sf(data = sf_map, size = 0.1, color = "darkgrey") +
  geom_sf(data = blackwood_acacia_sp, color = "darkgreen", size = 0.5) +
  theme_void()
```


Now and interactive map:
```{r}
tmap_mode("view")

tm_shape(blackwood_acacia_sp) +
  tm_dots()
```




