---
title: "ESM 244 - Lab 1"
author: "Kerry Nixon"
date: "1/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
```


### Read in SF trees data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Basic wrangling reminders

Refresh some skills for data wranginling and summary statistics using functions in the `dplyr` package.

find the top 5 highest observations of trees by legal_status, do some wrangling and make a graph.

count() can replace group_by summaraize. You can caount by multple variables seperated by commas. `count(variable_1, variable_2)`

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% # count by listed variable or variables
  drop_na(legal_status) %>% # remove rows with na in listed variable
  rename(tree_count = n) %>% # rename(new name = old name)
  relocate(tree_count) %>% # move tree_count to first column position
  slice_max(tree_count, n=5) # select the top 5 highest values
```













